# TODO: Lesson title here

{{< include ../includes/_wip.qmd >}}

```{r setup}
#| include: false
```

<!-- TODO: Add an introduction here -->

-   Continuous integration using GitHub Actions

**Session objectives**:

The overall objective for this session is to:
-   Explain what *automated testing* means when developing packages, 
    and deploy simple GitHub Actions to automate tests

More specific objectives are to:
-   Deploy a GitHub Action that runs CMD Check
-   Deploy a GitHub Action that runs test coverage

## TODO: Section

-   Continuous Integration and GitHub Actions

    -   Text: Explain what is continuous integration
    -   Text: Explain GitHub Actions
        (https://github.com/features/actions)\[https://github.com/features/actions\]
    -   Exercise: Deploy a sample GitHub Actions
        -   Probably one that install R using the r-lib action is a good
            exercise
    -   Code: `usethis::use_github_action_check_standard()`
    -   Text: Explain how to navigate the result of the GitHub Action
        and the logs window

::: callout-note
## Reading task: \~4 minutes

TODO: Reading text
:::

## Exercise: TODO: Add title

> Time: \## minutes.

```{r solution-LABEL}
#| eval: false
#| code-fold: true
#| code-summary: "**Click for the solution**. Only click if you are struggling or are out of time."
```

## Summary

-   TODO: List of summary items
