# Checking correctness of code: Debugging

-   Using basic debugging strategies ('breakpoints', simple `print()` or
    `stop()` calls, `debug()`, `browser()`)
-   Can take some from the intermediate course (I think that's where it
    is)

{{< include ../includes/_wip.qmd >}}

```{r setup}
#| include: false
```

<!-- TODO: Add an introduction here -->

**Session objectives**:

-   Enter the interactive debugger and use it to explore why unexpected
    behavior is occurring

## TODO: Section

-   Debugging your functions
    -   Text: Tracking down problems. Give code example with unexpected
        output. How do we figure out what the problem is?
    -   Code: Motivation? Maybe, something like:
        `> book_meta(c("Title: Dracula", "Author: Bram Stoker"))         # A tibble: 0 x 3         # … with 3 variables: author <chr>, title <chr>, release <chr>`
        Why don't we get a one row tibble back?
    -   Text: Print debugging. An easy way to start, but like a
        postmortem, can't step in and try to solve problem.
    -   Code: `print()`, `cat()`, `str()`
    -   Text: Diving inside your functions. Sometimes you want to try
        things inside your function. The interactive debugger puts you
        inside the function environment.
    -   Code: Adding Editor breakpoints
    -   Text: Navigating the interactive debugger, focus on Environment
        and Console, execute next statement, stop.
    -   Text: Other ways to enter the interactive debugger? e.g.
        `debug()`, `browser()`
    -   Exercise: Back to motivation, can you narrow down the problem
        and suggest a solution?
    -   Text: One solution,
        `extract_element("Author: Bram Stoker", "Title")`, would be more
        useful if returned `NA`. Use tests to describe bugs - test fails
        now, but you'll know you've fixed the bug once test passes.
    -   Code: Add tests that fail.
    -   Exercise: Fix the failing tests, and confirm all tests pass.

::: callout-note
## Reading task: \~4 minutes

TODO: Reading text
:::

## Exercise: TODO: Add title

> Time: \## minutes.

```{r solution-LABEL}
#| eval: false
#| code-fold: true
#| code-summary: "**Click for the solution**. Only click if you are struggling or are out of time."
```

## Summary

-   TODO: List of summary items
