# Checking correctness of code: Debugging

{{< include ../includes/_wip.qmd >}}

```{r setup}
#| include: false
```

**Introduction**

Debugging is one of activities that seems really scary and difficult,
but once you try it and use it, is not nearly as intimidating as it
seemed. To debug, which means to find and fix problems in your code,
there are several ways, e.g. simply adding 'breakpoints', `print()` or
`stop()` calls to your code, or by using the interactive debugger. All
these methods help you stop and/or inspect the code during its
execution, which is very helpful if you suspect there is a problem in
the code.

**Session objectives**:

-   Learn how to use RStudio's interactive debugger to explore
    unexpected behavior in your code.
-   Learn how and when to use `print()` and `stop()` calls to make
    debugging your code easier.
-   Learn how to interpret the `traceback()` error messages to find
    which function is causing the error *Too much/too advanced?*

## TODO: Section

### Debugging using the RStudio debugger

(Content is a copy/paste from intermediate course appendix C for now)

Of of the simplest way to debug is to insert the `browser()` function
into the the start of your function, re-running the function by either
manually running it or using source with {{< var keybind.source >}}, and
using it again.

For instance, we have a function like this, which returns the double of
the input minus the squareroot of the input:

```{r}
my_function <- function(number) {
  double_of_number <- number + number
  number_squareroot <- sqrt(number)
  return(double_of_number - number_squareroot)
}
```

When you use this function, it usually works without a problem, but some
times it returns `NaN` with a warning. Since this is a small toy
function with a helpful warning message, you may be able to spot the
location of the issue right away, but you may not always be so lucky.
The debugger can always help you, so let's use it to find out where the
problem arises in the function!

To start the debugger, you can either insert the `browser()` function
into the function itself:

```{r, eval=FALSE}
test_debugging <- function(number) {
  browser()
  double_of_number <- number + number
  number_squareroot <- sqrt(number)
  return(double_of_number - number_squareroot)
  
}
```

And re-run and use the function again, which will pop up a new debugging
panel in RStudio whenever the function is called.

Alternatively, you can start the debugger by wrapping a function (even a
function that you did not write) within a `debugonce()` call, which will
run that function in the debugger the next time it is called:

```{r, eval=FALSE}
debugonce(my_function)
my_function(2)
}
```

Sadly, we can't show this on the website since it only works in RStudio
(we may add a video as some point). When you are in the debugger will
open up and it will show a few things:

<!-- TODO: eventually add an image or video.. but for next year. -->

**Describe the debugger environment:**

-   What's different from the normal environment

-   How to use the new functionalities to step through a function,
    change variable values and inspect output.

-   Using 'step into current function call' in a more complicated
    function.

-   How to exit the debugger: 'Q'

-   A yellow line will highlight the code in the function, along with a
    green arrow on the left of the line number.

-   The Console will now start with `Browse[1]>` and will have text like
    `debug at ...`.

-   There will be new buttons on the top of the Console like "Next",
    "Continue", and "Stop".

-   The Environment pane will be empty and will say "Traceback".

<!-- TODO: Add image or video of this or extend it from previous video? -->

<!-- TODO: Image of debugging changes, with descriptions -->

In this mode you can really investigate what is happening with your code
and how to fix it. The way to figure out what's wrong is by running the
code bit by bit. This debug environment is empty except for the actions
that occur within it, so it really can help figure things out.

**From RSE**:

-   Debugging your functions
    -   Text: Tracking down problems. Give code example with unexpected
        output. How do we figure out what the problem is?
    -   Code: Motivation? Maybe, something like:
        `> book_meta(c("Title: Dracula", "Author: Bram Stoker"))         # A tibble: 0 x 3         # … with 3 variables: author <chr>, title <chr>, release <chr>`
        Why don't we get a one row tibble back?
    -   Text: Print debugging. An easy way to start, but like a
        postmortem, can't step in and try to solve problem.
    -   Code: `print()`, `cat()`, `str()`
    -   Text: Diving inside your functions. Sometimes you want to try
        things inside your function. The interactive debugger puts you
        inside the function environment.
    -   Code: Adding Editor breakpoints
    -   Text: Navigating the interactive debugger, focus on Environment
        and Console, execute next statement, stop.
    -   Text: Other ways to enter the interactive debugger? e.g.
        `debug()`, `browser()`
    -   Exercise: Back to motivation, can you narrow down the problem
        and suggest a solution?
    -   Text: One solution,
        `extract_element("Author: Bram Stoker", "Title")`, would be more
        useful if returned `NA`. Use tests to describe bugs - test fails
        now, but you'll know you've fixed the bug once test passes.
    -   Code: Add tests that fail.
    -   Exercise: Fix the failing tests, and confirm all tests pass.

::: callout-note
## Reading task: \~4 minutes

TODO: Reading text
:::

## Exercise: TODO: Add title

> Time: \## minutes.

```{r solution-LABEL}
#| eval: false
#| code-fold: true
#| code-summary: "**Click for the solution**. Only click if you are struggling or are out of time."
```

## Summary

-   TODO: List of summary items

## Resource for further reading

https://support.posit.co/hc/en-us/articles/205612627-Debugging-with-the-RStudio-IDE

